---
title: "Easy EDA in R"
author: "Barry DeCicco"
date: "6/10/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# library(tabyl)   tabyl is a function within janitor
library(janitor)
library(inspectdf)
library(skimr)

library(tidyverse)
```


```{r import_big_city_health_data, include=FALSE}
# df_health <- read.csv("https://query.data.world/s/t2obfytkhpqy5sygr4s4vftxjobmy6", header=TRUE, stringsAsFactors=FALSE);
```

```{r import_diamond_data, include=FALSE}
# df_diamonds <- read.csv("https://raw.githubusercontent.com/tidyverse/ggplot2/master/data-raw/diamonds.csv", header=TRUE, stringsAsFactors=FALSE);
```


```{r import_penguin_data, include=FALSE}
# df_penguins <- read.csv("https://raw.githubusercontent.com/allisonhorst/palmerpenguins/master/data-raw/penguins_raw.csv", header=TRUE, stringsAsFactors=FALSE);
```



```{r types_head_tail, echo = TRUE}

# head(df_penguins)
# tail(df_penguins)


```





## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

