---
title: "Easy EDA in R"
author: "Barry DeCicco"
date: "For Ann Arbor R Users' Group, June 11, 2020"
output: 
  html_document:
    css: style.css
    toc: true
    toc_float: true
    
# ioslides_presentation
---

# **What this presentation covers**
<br>
**This will not be a detailed coverage of EDA, but it will introduce some methods to make it easier. <br><br>This is important because hard EDA tends to be EDA not done.**
<br>
<br>

# **Why Do Exploratory Data Analysis?**
<br>
- **Make Sure that the import went well**
- **Find out what data you have - by column**
- **Find out what data you have - by row**

**The first point is not emphasized, but should be - when importing data sets, it's really easy to have things not be understood correctly by R, resulting in numeric/date variables being read in as strings, or missing value codes not being undersood.**



# **Load libraries:** 
<br>
- **gridExtra**
- **janitor** 
- **kableExtra** 
- **inspectdf**
- **skimr**
- **tidyverse**



```{r setup,  include=FALSE}  

# load libraries as needed:

library(gridExtra)
library(janitor)
library(kableExtra)
library(inspectdf)
library(skimr)

library(tidyverse)
```




# **Data - Penguins**
<br>
**This will use the 'penguins' data set, prepared by Allison Horst (https://github.com/allisonhorst):
<br>
https://raw.githubusercontent.com/allisonhorst/palmerpenguins/master/data-raw/penguins_raw.csv**
<br>

**Note that I have not worked with this data set before, so this is actual, gritty, real-life EDA!**
<br>
<br>


# **Import the data set** 
<br>
**Import the data set from a .csv file on Allison Horts' gitub site:**
<br>
```{r import_penguin_data, echo=TRUE}
df_penguins <- read.csv("https://raw.githubusercontent.com/allisonhorst/palmerpenguins/master/data-raw/penguins_raw.csv", header=TRUE, stringsAsFactors=FALSE);
```



# **Make Sure that the import went well**

- **Is the created object a data frame?**
- **What are the row/column counts? (what did you expect?)**
- **What are the column types?  (Numeric/Character/Dates as expected)**
<br>
<br>

```{r import_check_quality_1, echo=TRUE}
options(width = 80)

print(class(df_penguins))
print(dim(df_penguins))


```

# **Find the type of each column**

**This code produces a nice table:**

```{r import_check_quality_2, echo=TRUE}


# From:  https://stackoverflow.com/questions/40306280/how-to-transpose-a-dataframe-in-tidyverse

df_column_types <- lapply(df_penguins, class) %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value) %>%
  rename("Column_Name" = "name", "Column_Type" = "1")




gridExtra::grid.table(df_column_types)

```

# **Find out what data you have - by column**


- **List unique values and counts for each column**
- **List missing value counts**

#### This is where skimr comes in, rather than the standard summary commands available in R:


# **Skimr - character variables:**

```{r skimr_character, echo=TRUE }
options(width = 100)

 df_penguins %>%  
  dplyr::select_if(is.character) %>% 
  skim()



```



# **Skimr - numeric variables:**

```{r skimr_numeric, echo=TRUE}

df_penguins %>% 
  dplyr::select_if(is.numeric) %>% 
  skim()

```


# **Skimr handles grouped data:**

```{r skimr_numeric_grouped, echo=TRUE}

df_penguins %>%  
  dplyr::group_by(studyName) %>% 
  dplyr::select_if(is.numeric) %>% 
  skim()

```


# **Other Features:**

- **Behaves nicely in pipelines (i.e, feed results into other operations)**
- **Specify statistics and classes, using 'skim_with()**
- **skimr handles a variety of data types (e.g., character, factor, date,...)**
<br>
<br>
**See references for more details**


# **Inspectdf:**
<br>
**This has a number of functions. Here, we will just show the missing data plot function, using inspect_na():**

```{r inspectdf_missing, echo=TRUE}

df_penguins %>%  
  dplyr::select_if(is.numeric) %>% 
  inspect_na() %>% 
  show_plot()


```

# **DataExplorer**

**This packages has many features, and I will just show three:**

- **missing value plot:**
- **Boxplot matrix**
- **Corelation plot**
<br>
**Missing value plot:**
```{r DataExplorer_missing, echo=TRUE}

df_penguins %>%  
  dplyr::select_if(is.numeric) %>% 
  DataExplorer::plot_missing() 


```
<br>
**Boxplot:**
```{r DataExplorer_boxplot, echo=TRUE}

df_penguins %>%  
  DataExplorer::plot_boxplot(by='studyName')


```


<br>
**Correlation Plot:**
```{r DataExplorer_correlation_plot, echo=TRUE}

df_penguins %>% 
  dplyr::select_if(is.numeric) %>% 
  DataExplorer::plot_correlation(
    cor_args = list( 'use' = 'complete.obs')
    )


```

# **Conclusion:**

<br>
**These three packages will automate a lot of EDA functions, 
and offer the ability to customize and pipeline their functions.
They are superior to the standard R methods.**
<br>
<br>
<br>
<br>
<br>
<br>

# **Questions?**
<br>
<br>
<br>
<br>
<br>


# **References:**

<br>
**'rOpenSci: The skimr package' (https://docs.ropensci.org/skimr/index.html)**
<br>**'Display a Beautiful Summary Statistics in R using Skimr Package' (https://www.datanovia.com/en/blog/display-a-beautiful-summary-statistics-in-r-using-skimr-package/)**
<br>
**'inspectdf' (https://alastairrushworth.github.io/inspectdf/)**
<br>
**'Explore Your Dataset in R' (https://www.littlemissdata.com/blog/simple-eda)**
<br>
**'How to Automate EDA with DataExplorer in R' (https://www.programmingwithr.com/how-to-automate-eda-with-dataexplorer-in-r/)**
